Economic Evaluation of Mis-speficied Geographically Weighted Regression Models for Site-specific Nitrogen Management

### 2023-11-05 Writing/manuscript_gwr.rmd
addressed minor writing comments

The primary objective of this study is to examine the degree and direction of bias in the estimation of the economic value of GWR-based site-specific nitrogen rate recommendations (i.e., variable rate applications of nitrogen) when the yield response function assumed in the GWR model is specified as quadratic but is actually of quadratic-plateau functional form.  We will do so using Monte Carlo simulations, which, since we define and know the true yield response function in the data generating process, allow us to measure the degree of bias. Unlike in @griffith2008spatial, @harris2019simulation, @paez2011simulation, and @da2016multiple, our objective is not to measure the bias in the coefficient estimation by GWR, but rather to measure economic effects^[While the environmental impacts of fertilizer management are of great social importance, it is out of the scope of this study.] of the statistical inaccuracy in estimating the coefficients of the model.  We show that the misspecification tends to significantly over-estimate GWR's true economic value.

@@

The field was 432m (72 cells) wide and 864m (144 cells) long, making its area 37.3 ha. Data from the 6m-long headlands, side-lands (there are no transition zones on the north and south ends of the field), and 12m-wide "transition zones" were not included in the analysis. In total, data from 1440 subplots were available for experimental data analysis. The index $j \in {1, 2, \dots, 1440}$ identifies the field's subplots, and the index $i \in {1, \dots , 6}$ identifies the cells within a subplot. The couplet $(j, i)$ is used to identify individual cells.

@@

where $\textbf{c}_{j,i} = (\alpha_{j,i},\beta_{j,i},\gamma_{j,i},\tau_{j,i})$ is the vector of yield response function parameters. $\tau_{j,i}$ can be interpreted as the critical value of N rate at which cell $(j, i)$'s yield reaches its plateau, where yield no longer increases with N rate. $\varepsilon_{j, i}$ is a stochastic error term. The quadratic-plateau functional form implies a diminishing marginal product of N until a yield plateau, which are desirable properties for a yield response function [@frank1990comparison] and a pattern commonly observed in agronomic studies. Note that the cell-specific yield response function $f_{j, i} (N)$ is the reduced form of the meta response function $f$ when the characteristics vector takes on the value $\textbf{c}_{j, i}$. In other words, all the soil or field properties that could affect yield response to N are embedded in the simulated parameters.

>>>>

The primary objective of this study is to examine the degree and direction of bias in the estimation of the economic value of GWR-based site-specific nitrogen rate recommendations (i.e., variable rate applications of nitrogen) when the yield response function assumed in the GWR model is specified as quadratic but is actually of quadratic-plateau functional form.  We will do so using Monte Carlo simulations, which, since we define and know the true yield response function in the data generating process, allow us to measure the degree of bias. Unlike in @griffith2008spatial, @harris2019simulation, @paez2011simulation, and @da2016multiple, our objective is not to measure the bias in the coefficient estimation by GWR, but rather to measure economic effects^[While the environmental impacts of fertilizer management are of great social importance, it is out of the scope of this study. The "economically optimal" in the context of this study does not mean "socially optimal", but is rather limited to the economic optimization of farmers' private production.] of the statistical inaccuracy in estimating the coefficients of the model.  We show that the misspecification tends to significantly over-estimate GWR's true economic value.

@@

The field was 432m (72 cells) wide and 864m (144 cells) long, making its area 37.3 ha. Data from the 6m-long headlands, side-lands (there are no transition zones on the north and south ends of the field), and 12m-wide "transition zones" were not included in the analysis. In total, simulated data from 1440 subplots were used for data analysis. The index $j \in {1, 2, \dots, 1440}$ identifies the field's subplots, and the index $i \in {1, \dots , 6}$ identifies the cells within a subplot. The couplet $(j, i)$ is used to identify individual cells.

@@

where $\textbf{c}_{j,i} = (\alpha_{j,i},\beta_{j,i},\gamma_{j,i},\tau_{j,i})$ is the vector of yield response function parameters. $\tau_{j,i}$ can be interpreted as the critical value of N rate at which cell $(j, i)$'s yield reaches its plateau, where yield no longer increases with N rate. $\varepsilon_{j, i}$ is a stochastic error term following a normal distribution and is spatially autocorrelated. The quadratic-plateau functional form implies a diminishing marginal product of N until a yield plateau, which are desirable properties for a yield response function [@frank1990comparison] and a pattern commonly observed in agronomic studies. Note that the cell-specific yield response function $f_{j, i} (N)$ is the reduced form of the meta response function $f$ when the characteristics vector takes on the value $\textbf{c}_{j, i}$. In other words, all the soil or field properties that could affect yield response to N are embedded in the simulated parameters.
